---
title: "hw01-Caroline-Seip"
author: "Caroline"
date: "September 16, 2018"
output: html_document
---

In this document we will test some basic functions of R that can be used to explore a dataset.

# Loading the data

Start by loading in the 'gapminder' dataset, which we will be exploring. If you don't already have this dataset downloaded, run 'install.packages("gapminder") in your R console.
```{r}
library(gapminder)
```
Now the gapminder dataset is loaded.

# Data exploration

Now we will start to explore the dataset, first let's look at the first few rows of data that the dataset contains using the 'head' function:
```{r}
head(gapminder)
```
We can see here that there are columns for country, continent, year, life expectancy, population, and per capita GDP. The head function shows us the first six rows of data. 

We can also take a look at the structure of the dataset using the 'str' function:

```{r}
str(gapminder)
```
1704 observations of 6 variables

- country is a factor with 142 levels
- continent is a factor with 5 levels
- year is an integer, beginning in 1952 with increments of 5 years
- lifeExp is the life expectancy for each country in years
- pop is the population of each country in a specific year
- gdpPercap is the per capita GDP in each country in a specific year

We can also just determine the number of columns in the dataset using the 'ncol' function:

```{r}
ncol(gapminder)
```
The dataset has 6 columns.

# Descriptive statistics

We can look at the mean, median, maximum and minimum of a variable in the dataset using these functions:
```{r}
mean(gapminder$lifeExp)
median(gapminder$lifeExp)
min(gapminder$lifeExp)
max(gapminder$lifeExp)
```
Life expectancy (years)
  Mean: 59.5
  Median: 60.7
  Minimum: 23.6
  Maximum: 82.6




